### 2022-01-11 paper/paper.tex
tune figures to keep page limit
Apply suggestions from code review
Co-authored-by: Michael Schlottke-Lakemper <michael@sloede.com>

\begin{figure*}[htbp]
  \includegraphics[width=\linewidth]{../figures/trixi_global_overview}
  \caption{Schematic overview of the basic components in \trixi and how they
           interact.}
  \label{fig:trixi_global_overview}
\end{figure*}

@@

\begin{figure}[!h]
  \includegraphics[width=\linewidth]{../figures/pid_euler_mesh}
  \caption{Initial density of the compressible Euler problem on a slice of the
           periodic, curved mesh used for the performance benchmarks.}
  \label{fig:pid_euler_mesh}
\end{figure}

@@

While our Julia code is faster than the mature HPC Fortran code FLUXO for this non-trivial
example, we do not claim that Julia is generally faster than Fortran, C, or C++.
Instead, we would like to emphasize that well-written Julia code can be \emph{at least
as fast} as code written in these traditional scientific computing languages,
as demonstrated also by several microbenchmarks \cite{bezanson2017julia}.
Further, we expect to be able to achieve similar performance with either language
by spending enough time and effort to optimize the respective codes.
\trixi owes its performance optimizations in part to the code introspection and
profiling tools available in Julia. Similar tools used to optimize the performance in other
languages are often not as easy to use as their Julia counterparts. Therefore,
we developed some additional performance improvements in \trixi and ported them
to FLUXO later \cite{ranocha2021efficient}. Moreover, \trixi uses slightly
different implementations of computationally intensive kernels and stores a bit
less information in RAM, recomputing some terms instead. Additionally, the
manufactured solution of the compressible Euler equations uses source terms
containing sine and cosine terms, which are computed together by \lstinline{sincos}
in \trixi but require individual calls to \lstinline{sin} and \lstinline{cos} in
FLUXO.

@@

As demonstrated in Section~\ref{sec:performance-comparison},
Julia is not generically slower than traditional high-performance programming
languages like C, C++, and Fortran. A minor exception in the context of scientific
computing relying on floating point operations is that Julia does not perform
automatic fused multiply-add (FMA) contraction, that is replacing a multiplication
followed by an addition with a single FMA instruction, as it is inconsistent
with its strict floating point semantics. In other compilers, this optimization
is controlled by a global compiler option (\lstinline{-ffp-contract} in GCC and
Clang) and often enabled at reasonable optimization levels such as \lstinline{-O2}
for GCC and Intel compilers\footnote{An interesting coincidence is that not only
Julia but most LLVM-based frontends tend not to enable FMA contraction by default.}.
In Julia, FMA contraction is controlled more locally: the
\lstinline{muladd}\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}
intrinsic function allows the compiler to evaluate a multiply-add using the most
efficient method. Relying on Julia's code manipulation techniques, the
\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}
macro makes it convenient to use the \lstinline{muladd} function by
syntactically rewriting expressions to insert \lstinline{muladd} in appropriate
locations.

>>>>

\begin{figure*}[htbp]
\centering
  \includegraphics[width=0.9\linewidth]{../figures/trixi_global_overview}
  \caption{Schematic overview of the basic components in \trixi and how they
           interact.}
  \label{fig:trixi_global_overview}
\end{figure*}

@@

\begin{figure}[!h]
\centering
  \includegraphics[width=0.9\linewidth]{../figures/pid_euler_mesh}
  \caption{Initial density of the compressible Euler problem on a slice of the
           periodic, curved mesh used for the performance benchmarks.}
  \label{fig:pid_euler_mesh}
\end{figure}

@@

While our Julia code is faster than the mature HPC Fortran code FLUXO for this non-trivial
example, we do not claim that Julia is generally faster than Fortran, C, or C++.
Instead, we would like to emphasize that well-written Julia code can be \emph{at least
as fast} as code written in these traditional scientific computing languages,
as demonstrated also by several microbenchmarks \cite{bezanson2017julia}.
Further, we expect to be able to achieve similar performance with either language
by spending enough time and effort to optimize the respective codes.
\trixi owes its performance optimizations in part to the code introspection and
profiling tools available in Julia. Similar tools used to optimize the performance in other
languages are often not as easy to use as their Julia counterparts. Therefore,
we developed some additional performance improvements in \trixi and ported them
to FLUXO later \cite{ranocha2021efficient}. Moreover, \trixi uses slightly
different implementations of computationally intensive kernels and stores marginally
less information in memory, recomputing some terms instead. Additionally, the
manufactured solution of the compressible Euler equations uses source terms
containing sine and cosine terms, which are computed together by \lstinline{sincos}
in \trixi but require individual calls to \lstinline{sin} and \lstinline{cos} in
FLUXO.

@@

As demonstrated in Section~\ref{sec:performance-comparison},
Julia is not generically slower than traditional high-performance programming
languages like C, C++, and Fortran. A minor exception in the context of scientific
computing relying on floating point operations is that Julia does not perform
automatic fused multiply-add (FMA) contraction, that is, replacing a multiplication
followed by an addition with a single FMA instruction, as it is inconsistent
with its strict floating point semantics. In other compilers, this optimization
is controlled by a global compiler option (\lstinline{-ffp-contract} in GCC and
Clang) and often enabled at reasonable optimization levels such as \lstinline{-O2}
for GCC and Intel compilers\footnote{An interesting coincidence is that not only
Julia but most LLVM-based frontends tend not to enable FMA contraction by default.}.
In Julia, FMA contraction is controlled more locally: the
\lstinline{muladd}\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}
intrinsic function allows the compiler to evaluate a multiply-add using the most
efficient method. Relying on Julia's code manipulation techniques, the
\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}
macro makes it convenient to use the \lstinline{muladd} function by
syntactically rewriting expressions to insert \lstinline{muladd} in appropriate
locations.
